Hazelnut: A Bidirectionally Typed \\ Structure Editor Calculus

### 2016-07-06 formally.tex

The rules above are numerous and fairly uninteresting. That makes them quite hazardous -- we might make a mistake absent-mindedly. One check against this is to prove a lemma that establishes that movement actions move the cursor but do not change the cursor erasure.
\begin{lemma}[Movement Erasure Invariance]\label{lemma:movement-erasure} \todo{make lemmas number with theorems} 
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\performMove{\ztau}{\aMove{\delta}}{\ztau'}$ then $\removeSel{\ztau}=\removeSel{\ztau'}$.
\item If $\performMove{\zexp}{\aMove{\delta}}{\zexp'}$ then $\removeSel{\zexp}=\removeSel{\zexp'}$.
\end{enumerate}
\end{lemma}
\noindent Lemma \ref{lemma:movement-erasure} is useful also in that the relevant cases of Theorem \ref{thrm:actsafe} are straightforward by its application.

@@

@@

\begin{theorem}[Reachability] ~
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\removeSel{\ztau}=\removeSel{\ztau'}$ then there exists some $\bar\alpha$ such that $\bar{\alpha}~\mathsf{movements}$ and $\performTyp{\ztau}{\bar\alpha}{\ztau'}$.
\item If $\hana{\hGamma}{\removeSel{\zexp}}{\htau}$ and $\removeSel{\zexp}=\removeSel{\zexp'}$ then there exists some $\bar{\alpha}$ such that $\bar{\alpha}~\mathsf{movements}$ and $\performAna{\hGamma}{\zexp}{\htau}{\bar{\alpha}}{\zexp'}$. 
\item If $\hsyn{\hGamma}{\removeSel{\zexp}}{\htau}$ and $\removeSel{\zexp}=\removeSel{\zexp'}$ then there exists some $\bar{\alpha}$ such that $\bar{\alpha}~\mathsf{movements}$ and $\performSyn{\hGamma}{\zexp}{\htau}{\bar\alpha}{\zexp'}{\htau}$.
\end{enumerate}
\end{theorem}

To prove Reachability, we break it into two lemmas. One...\todo{write this}

>>>>

The rules above are numerous and fairly uninteresting. That makes them quite hazardous -- we might make a mistake absent-mindedly. One check against this is to prove a lemma that establishes that movement actions move the cursor but do not change the cursor erasure.
\begin{lemma}[Movement Erasure Invariance]\label{lemma:movement-erasure}
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\performMove{\ztau}{\aMove{\delta}}{\ztau'}$ then $\removeSel{\ztau}=\removeSel{\ztau'}$.
\item If $\performMove{\zexp}{\aMove{\delta}}{\zexp'}$ then $\removeSel{\zexp}=\removeSel{\zexp'}$.
\end{enumerate}
\end{lemma}
\noindent Lemma \ref{lemma:movement-erasure} is useful also in that the relevant cases of Theorem \ref{thrm:actsafe} are straightforward by its application.

@@

A simple auxiliary judgement, $\bar\alpha~\mathsf{movements}$, defined in the supplement, establishes that $\bar\alpha$ consists only of actions of the form $\aMove{\delta}$.\todo{define this in supplement}

@@

\begin{theorem}[Reachability]\label{thrm:reachability} ~
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\removeSel{\ztau}=\removeSel{\ztau'}$ then there exists some $\bar\alpha$ such that $\bar{\alpha}~\mathsf{movements}$ and $\performTyp{\ztau}{\bar\alpha}{\ztau'}$.
\item If $\hana{\hGamma}{\removeSel{\zexp}}{\htau}$ and $\removeSel{\zexp}=\removeSel{\zexp'}$ then there exists some $\bar{\alpha}$ such that $\bar{\alpha}~\mathsf{movements}$ and $\performAna{\hGamma}{\zexp}{\htau}{\bar{\alpha}}{\zexp'}$. 
\item If $\hsyn{\hGamma}{\removeSel{\zexp}}{\htau}$ and $\removeSel{\zexp}=\removeSel{\zexp'}$ then there exists some $\bar{\alpha}$ such that $\bar{\alpha}~\mathsf{movements}$ and $\performSyn{\hGamma}{\zexp}{\htau}{\bar\alpha}{\zexp'}{\htau}$.
\end{enumerate}
\end{theorem}

The simplest way to prove Theorem \ref{thrm:reachability} is to  break it into two lemmas. Lemma \ref{lemma:reach-up} establishes that you can always move the cursor to the outermost position in an expression. This serves as a check on our $\aMove{\dParent}$ rules.
\begin{lemma}[Reach Up]\label{lemma:reach-up} ~
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\removeSel{\ztau}=\htau$ then there exists some $\bar\alpha$ such that $\bar\alpha~\mathsf{movements}$ and $\performTyp{\ztau}{\bar\alpha}{\zwsel{\htau}}$.
\item If $\hana{\hGamma}{\hexp}{\htau}$ and $\removeSel{\zexp}=\hexp$ then there exists some $\bar\alpha$ such that $\bar\alpha~\mathsf{movements}$ and $\performAna{\hGamma}{\zexp}{\htau}{\bar\alpha}{\zwsel{\hexp}}$. 
\item If $\hsyn{\hGamma}{\hexp}{\htau}$ and $\removeSel{\zexp}=\hexp$ then there exists some $\bar\alpha$ such that $\bar\alpha~\mathsf{movements}$ and $\performSyn{\hGamma}{\zexp}{\htau}{\bar\alpha}{\zwsel{\hexp}}{\htau}$.
\end{enumerate}
\end{lemma}
Lemma \ref{lemma:reach-down} establishes that you can always move the cursor from the outermost position to any other position. This serves as a check on our $\aMove{\dChild}$ and $\aMove{\dNext}$ rules. 
\begin{lemma}[Reach Down]\label{lemma:reach-down} ~
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\removeSel{\ztau}=\htau$ then there exists some $\bar\alpha$ such that $\bar\alpha~\mathsf{movements}$ and $\performTyp{\zwsel{\htau}}{\bar\alpha}{\ztau}$.
\item If $\hana{\hGamma}{\hexp}{\htau}$ and $\removeSel{\zexp}=\hexp$ then there exists some $\bar\alpha$ such that $\bar\alpha~\mathsf{movements}$ and $\performAna{\hGamma}{\zwsel{\hexp}}{\htau}{\bar\alpha}{\zexp}$.
\item If $\hsyn{\hGamma}{\hexp}{\htau}$ and $\removeSel{\zexp}=\hexp$ then there exists some $\bar\alpha$ such that $\bar\alpha~\mathsf{movements}$ nand $\performSyn{\hGamma}{\zwsel{\hexp}}{\htau}{\bar\alpha}{\zexp}{\htau}$.
\end{enumerate}
\end{lemma}
Theorem \ref{thrm:reachability} follows by straightforward composition of these two lemmas. (The witness to the existential is not, of course, the shortest path between any two cursor locations.)
